R-LIME\@: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-07-10 paper.tex
ICPR really final revised version

\begin{document}
\begin{center}
  % smaller title font only for rebuttal
  % additional two empty lines at the end of the title
  \vspace*{-22pt}
  \begin{tabular}[t]{c}{
      \large\bf
      Paper ID \paperID{}
    }
  \end{tabular}
  \par
  % additional small space at the end of the author name
  % additional empty line at the end of the title block
\end{center}
We deeply appriciate all the reviewers who made sharp and valuable comments
on our paper and gave us the opportunity to submit a revised version.
We have done our best for the revision according to their comments as shown below.
We also would like to thank in advance all the reviewers for reconsidering
the suitability of our revised paper for the conference.

\section*{Summary of revisions}
We would like to appriciate all the reviewers for their clear and insightful
comments.
We improved our paper from the original version according to their comments.
The major revision points are listed as follows:
\begin{enumerate}
  \item We added an overview of other existing methods in \cref{sec:related-work}.
  \item We added discussion about parameter selection in \cref{sec:param}.
  \item We conducted qualitative and quantitative comparisons
        between Anchor and R-LIME in \cref{sec:exp-qual,sec:exp-anchor}.
\end{enumerate}
Please see below for a point-by-point response to each reviewer's comments.

@@

        \begin{mycomment}
          We would appreciate for your suggestion.
          We added \cref{sec:param} as a guidance on how parameters should be
          selected in practice.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your valuable suggestion.
          We should add an overview of other model-agnostic interpretability
          methods, but we only added names of some other methods to
          \cref{sec:related-work}, because of lack of space.
        \end{mycomment}

@@

        \begin{mycomment}
          We would like to appriciate your sharp observations.
          We additionally conducted qualitative and quantitative comparisons
          between Anchor and R-LIME,
          which demonstrated that R-LIME generates more informative and
          general explanations than Anchor (see \cref{sec:exp-qual,sec:exp-anchor}).
        \end{mycomment}
\end{enumerate}
\section*{Responding to Reviewer \#4}
The authors should kindly explain why there is not a quantitative comparison
between their proposed method R-LIME and Anchor. The lack of the comparison
makes it hard to evaluate the impact of the proposed model.

\begin{mycomment}
  We would appreciate for your insightful comments.
  We additionally conducted qualitative and quantitative comparisons
  between Anchor and R-LIME,
  which demonstrated that R-LIME generates more informative and
  general explanations than Anchor (see \cref{sec:exp-qual,sec:exp-anchor}).
\end{mycomment}

@@

        \begin{mycomment}
          We would like to appriciate for your insightful suggestion.
          We interpreted your suggestrion to mean that we should move
          \cref{fig:example} and its explanation to \cref{sec:related-work}.
          We discussed about the most understandable structure of our paper.
          And for clarity, we concluded that we should present effectiveness
          of R-LIME compared to existing methods closely related to our method
          in \cref{sec:introduction},
          before we overview other existing methods in \cref{sec:related-work}.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your valuable feedback.
          It is true that we must address theoretical issues for using
          KL-LUCB algorithm, although our results suggest that its practical
          effect is negligible.
          We focus on presenting our novel framework for local interpretability,
          not on the detailed algorithm.
          Thus, we leave the issues for future work because of lack of space.
          But we deeply appreciate for your sharp observations.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your suggestion. We renamed the section to `Discussion'
          for clearly reflecting the content.
        \end{mycomment}
\end{itemize}

@@

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
\begin{abstract}
  In recent years,
  complex machine learning models have been introduced
  in various industrial fields due to their high accuracy.
  However,
  their increasing complexity has been a major obstacle to implementation
  in sensitive decision-making situations.
  In order to address this problem,
  various post-hoc explanation methods have been proposed,
  but they have not been able to achieve interpretability of
  both the explanation and its scope.
  We propose a new method, R-LIME,
  which interprets a complex classifier in an interpretable scope.
  R-LIME locally and linearly approximates a complex decision boundary
  of a black-box classifier in a rectangular region
  and maximizes the region as long as the approximation accuracy
  exceeds a given threshold.
  The resulting rectangular region is interpretable for users because it is
  expressed as a conjunction of feature predicates.
  Through qualitative and quantitative comparisons with the existing method
  on a real-world dataset,
  we demonstrate that R-LIME provides more reliable and interpretable
  explanations than existing methods.
  % 153 words
  \keywords{Interpretable machine learning \and Local surrogate model}
\end{abstract}

\section{Introduction}\label{sec:introduction}
In recent years,
complex machine learning models, such as deep neural networks and random forests,
have been widely introduced in various industrial fields
due to their significant improvement in accuracy.
However,
their increasing complexity and black-box nature pose challenges,
particularly in critical decision-making scenarios such as healthcare and finance,
where the opacity of decision process becomes a major obstacle to implementation.
In order to address this problem,
there has been extensive research in the field of post-hoc explanations
for machine learning models
~\cite{guidotti2018local,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure,
and the latter are further classified into \emph{global} and \emph{local} methods
based on their locality in input space~\cite{samek2021explaining}.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \vspace{-2.24cm}
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method)
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

@@

To address these limitations,
we propose a new method called R-LIME (Ruled LIME),
which provides both the contributions of each feature to the output
and the effective scope of the explanation.
R-LIME linearly approximates a complex decision boundary
in a rectangular region and maximizes the region
as long as the accuracy of the linear classifier exceeds a given threshold.
The region is interpretable for users because it is
expressed as a conjunction of feature predicates.
An example of the explanation by R-LIME for a sentiment prediction model
is shown in \cref{fig:example-rlime}.
It is clear that users can apply the insights derived from the explanation
only to the sentences containing the word ``not''.

@@

They are broadly divided into
\emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Model-dependent methods,
such as deep Taylor decomposition (DTD)~\cite{montavon2017explaining}
and layer-wise relevance propagation (LRP)~\cite{bach2015pixel},
most of which focus on neural networks and
explain the model's behavior using its parameters~\cite{samek2021explaining}.
While these methods provide detailed explanations
(e.g., layer-wise explanations for neural networks),
it is often challenging
to apply the same method to models with different structures.
In contrast,
model-agnostic methods use only the model's output.
Although they are applicable to any model,
they cannot explain the reasoning process inside the model.

Furthermore, model-agnostic methods are categorized into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as partial dependence plots (PDP)~\cite{friedman2001greedy}
and accumulated local effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as \hl{indivisual conditional expectation (ICE)}~\cite{goldstein2015peeking},
local interpretable model-agnostic explanations
(LIME)~\cite{ribeiro2016why}, Anchor~\cite{ribeiro2018anchors} and
\hl{shapley additive explanations (SHAP)}~\cite{lundberg2017unified},
explain model's behavior in the vicinity of a specific input.
While they offer explanations more simple and accurate
than global methods,
the scope of the explanation is limited locally.

@@

\subsection{Previous Work}
We specifically focus on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation is performed by the following steps:
\begin{enumerate}
  \item Generating a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) $x$ is converted into a binary vector
        $x'\in{\{0,1\}}^{m'}$,
        (ii) perturbed samples are generated by drawing non-zero elements
        from $x'$ uniformly at random,
        and (iii) the perturbed samples are converted back to the original space.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically defined using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, it does not explicitly indicate the region for generating perturbed samples,
making it difficult for users to assess the effective scope of the explanation
~\cite{ribeiro2018anchors}.

@@

\subsection{Overview}
We propose R-LIME,
a method that aims to address the limitations of LIME~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors}.
Our method locally approximates the given black-box classifier $f$
around the focal point $x$ by a linear classifier $g$ similar to LIME,
but generates the perturbed samples
from a rectangular region similar to Anchor
so that the generality of approximation is explicitly provided
(\cref{fig:rlime}).

@@

  \begin{algorithm}[p]
    \small
    \input{src/tex/alg4}
  \end{algorithm}
}
The algorithm of R-LIME
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main-problem},
greedy search are often used.
But greedy methods tend to converge to local optima,
so we use beam search, which selects multiple candidates at each iteration.
The pseudo-code is shown in \cref{alg:greedy-search}.

@@

\subsubsection{Searching Rules with Highest Accuracy}
Given the set of candidate rules $\cands$,
the algorithm selects the $B$ candidate rules with the highest accuracy.
This problem can be formulated
as best arm identification in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered as an arm,
and reward of arm $a_i$ follows a Bernoulli distribution
with $P(X=1)=\Prec(A_i)$.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using $z$ and $f(z)$ after each trial.
To efficiently search the rule (arm) with the highest accuracy,
we employ KL-LUCB algorithm~\cite{kaufmann2013information}.
The pseudo-code is shown in \cref{alg:best-cands}.
For tolerance $\epsilon\in[0,1]$, KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy the assumption.
This issue is discussed further in \cref{sec:reward}.

@@

\subsection{Computational Complexity}\label{subsec:complexity}
Post-hoc explanation methods including LIME, Anchor, and R-LIME
need to sample a perturbation vector
and get the output of the black-box model multiple times,
which is computationally expensive.
The number of samples required for LIME is $|\mathcal{Z}_p|$,
which is the number of samples designated by the user.
On the other hand,
the expected number of samples required for Anchor and R-LIME is bounded by
$\mathcal{O}[m\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}]$,
where $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$
is the expected number of samples
for best arm identification finding the best $B$ arms from $B\cdot m$ arms.
For KL-LUCB algorithm~\cite{kaufmann2013information},
\begin{equation}
  \mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}=
  \mathcal{O}\left[\frac{Bm}{\epsilon^2}\log\frac{Bm}{\epsilon^2\delta}\right].
\end{equation}
Then the total expected number of samples for Anchor and R-LIME is bounded by
\begin{equation}
  \mathcal{O}\left[\frac{Bm^2}{\epsilon^2}\log\frac{Bm}{\epsilon^2\delta}\right].
\end{equation}

For each iteration of KL-LUCB algorithm, R-LIME needs to update
the linear classifier $g_i$, which is not required in Anchor.
If we use logistic regression as the linear classifier and update it
by stochastic gradient descent (SGD)~\cite{robbins1951stochastic},
the computational complexity of updating $g_i$ is $\mathcal{O}(m)$.
It is negligible compared to the complexity of generating a perturbed sample,
which is $\mathcal{O}(m^2)$ if we get a sample
from a multivariate normal distribution using Cholesky decomposition in advance.
Overall, the computational complexity of R-LIME is comparable to that of Anchor.

\section{Experiments}
To verify the effectiveness of our method,
we compared LIME, \hl{Anchor} and R-LIME using a real-world dataset.
\hl{Our code for R-LIME is available on GitHub}\footnote{\url{https://github.com/g-ohara/rlime}}.

@@

We used the recidivism dataset~\cite{schmidt1988predicting} for our experiments.
The dataset contains personal information on 9549 prisoners released from
North Carolina prisons between July 1, 1979, and June 30, 1980.
As shown in \cref{tab:rcdv}, the dataset includes 19 items such as
race (\emph{Race}), gender (\emph{Gender}),
presence of alcohol dependence (\emph{Alcohol}),
number of prior offenses (\emph{Priors}),
and presence of recidivism (\emph{Recidivism}).
For this experiment,
we treated the binary classification problem of predicting
the presence of recidivism (\emph{Recidivism}) as the target label.
We discretized continuous features and removed missing values,
resulting in 15 features.

We splitted the dataset into training data (7639 instances) and test data (955 instances),
and trained a random forest model with 50 trees as the black-box classifier
using the training data.
Then, we generate LIME, \hl{Anchor} and R-LIME explanations
for two instances extracted from the test data (\cref{fig:instance}).
For R-LIME, we used logistic regression as the linear approximation model,
and a multivariate normal distribution estimated from the training data
as the distribution $\mathcal{D}$.
\hl{For both Anchor and R-LIME,} the beam width was set to $B=10$,
the confidence coefficient to $1-\delta=0.95$,
and the tolerance of KL-LUCB algorithm to $\epsilon=0.05$.
The accuracy threshold $\tau$ was set to $\tau=0.70,0.90$.

@@

\subsection{\hl{Quantitative Evaluation: Anchor vs. R-LIME}}\label{sec:exp-anchor}
\subsubsection{\hl{Experimental Setup}}
\hl{%
  To demonstrate that R-LIME explanations are more general than Anchor,
  we conducted a comparison of the coverage of explanations
  between Anchor and R-LIME\@.
  Under the same settings as in
}\cref{sec:exp-setting},
\hl{%
  we generated Anchor and R-LIME explanations
  for 100 instances from the test data of the recidivism dataset,
  under the values of $\tau=0.65,0.70,0.75,0.80,0.85,0.90$.
}

@@

\section{Conclusion}
Existing methods for
local model-agnostic explanations of black-box classifiers,
such as LIME and Anchor,
have limitations that they cannot achieve interpretability of
both the explanation and its application scope.
To address these challenges,
we proposed R-LIME,
a method that locally and linearly approximates the decision boundary
of a black-box classifier and provides a rectangular approximation region,
which is interpretable for users due to being expressed as a conjunction of feature predicates.
We proposed an algorithm to
maximize coverage of the approximation region
as long as the accuracy of the linear approximation model exceeds a given threshold.
Comparing R-LIME with existing methods on the real-world dataset,
we demonstrated that R-LIME achieves interpretability of both the explanation
and its application scope,
and provides explanations more accurate than LIME and more general than Anchor.
Finally,
we discussed the instability of behavior against imbalanced label distributions,
raised questions about the theoretical validity of using KL-LUCB algorithm,
and hyperparameter tuning in practice.

>>>>

\begin{document}
\begin{center}
  % smaller title font only for rebuttal
  % additional two empty lines at the end of the title
  \vspace*{-22pt}
  \begin{tabular}[t]{c}{
      \large\bf
      Paper ID \paperID{}
    }
  \end{tabular}
  \par
  % additional small space at the end of the author name
  % additional empty line at the end of the title block
\end{center}
We deeply appreciate all the reviewers who made sharp and valuable comments
on our paper and gave us the opportunity to submit a revised version.
We have done our best for the revision according to their comments as shown below.
We also thank in advance all the reviewers for reconsidering
our revised paper's suitability for the conference.

\section*{Summary of revisions}
We appreciate all the reviewers for their clear and insightful
comments.
We have improved our paper based on their comments.
The major revision points are listed as follows:
\begin{itemize}
  \item We mentioned the names of other existing methods in \cref{sec:related-work}.
  \item We added discussion about parameter selection in \cref{sec:param}.
  \item We conducted qualitative and quantitative comparisons
        between Anchor and R-LIME in \cref{sec:exp-qual,sec:exp-anchor}.
  \item We renamed the section `Challenges and Future Work' to `Discussion'.
\end{itemize}
Please see below for a point-by-point response to each reviewer's comments.

@@

        \begin{mycomment}
          We appreciate your suggestion.
          We added \cref{sec:param} to provide guidance on
          how parameters should be selected in practice.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your valuable suggestion.
          Although we acknowledge the importance of including an overview of
          other model-agnostic interpretability methods,
          due to space constraints, we only mentioned the names of some
          other methods in \cref{sec:related-work}.
        \end{mycomment}

@@

        \begin{mycomment}
          We appreciate your sharp observations.
          We additionally conducted qualitative and quantitative comparisons
          between Anchor and R-LIME,
          which demonstrated that R-LIME generates more informative and
          general explanations than Anchor (see \cref{sec:exp-qual,sec:exp-anchor}).
        \end{mycomment}
\end{enumerate}
\section*{Responding to Reviewer \#4}
The authors should kindly explain why there is not a quantitative comparison
between their proposed method R-LIME and Anchor. The lack of the comparison
makes it hard to evaluate the impact of the proposed model.

\begin{mycomment}
  We appreciate your insightful comments.
  We additionally conducted qualitative and quantitative comparisons
  between Anchor and R-LIME,
  which demonstrated that R-LIME generates more informative and
  general explanations than Anchor (see \cref{sec:exp-qual,sec:exp-anchor}).
\end{mycomment}

@@

        \begin{mycomment}
          We appreciate your insightful suggestion.
          We interpreted your suggestion to mean that we should move
          \cref{fig:example} and its explanation to \cref{sec:related-work}.
          After discussing the most understandable structure of our paper,
          we concluded that we should present effectiveness of R-LIME
          compared to closely related methods in \cref{sec:introduction},
          before we overview other existing methods in \cref{sec:related-work}.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your valuable feedback.
          While our results suggest that the practical effect of using
          the KL-LUCB algorithm is negligible,
          we acknowledge the need to address its theoretical issues.
          Our focus is on presenting our novel framework for local
          interpretability, not the detailed algorithm.
          Thus, we leave the issues for future work due to space constraints.
          However, we deeply appreciate your sharp observations.
        \end{mycomment}

@@

        \begin{mycomment}
          Thank you for your suggestion. We renamed the section to `Discussion'
          to clearly reflect the content.
        \end{mycomment}
\end{itemize}

@@

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
\begin{abstract}
  In recent years,
  complex machine learning models have been widely introduced
  in various industrial fields due to their high accuracy.
  However,
  their increasing complexity has been a major obstacle to their implementation
  in sensitive decision-making situations.
  In order to address this problem,
  various post-hoc explanation methods have been proposed,
  but they have not been able to achieve interpretability of
  both the explanation and its scope.
  We propose R-LIME, a novel method that interprets complex classifiers
  within an interpretable scope.
  R-LIME locally and linearly approximates the complex decision boundary
  of a black-box classifier within a rectangular region
  and maximizes the region as long as the approximation accuracy
  exceeds a given threshold.
  The resulting rectangular region is interpretable for users because it is
  expressed as a conjunction of feature predicates.
  Through qualitative and quantitative comparisons on a real-world dataset,
  we demonstrate that R-LIME provides more reliable and interpretable
  explanations than existing methods.
  % 153 words
  \keywords{Interpretable machine learning \and Local surrogate model}
\end{abstract}

\section{Introduction}\label{sec:introduction}
In recent years,
complex machine learning models, such as deep neural networks and random forests,
have been widely introduced in various industrial fields
due to their significant improvements in accuracy.
However,
their increasing complexity and black-box nature pose challenges,
particularly in critical decision-making scenarios such as healthcare and finance,
where the lack of transparency becomes a major obstacle to implementation.
In order to address this problem,
extensive research has focused on post-hoc explanations for complex models
~\cite{guidotti2018local,ribeiro2016why,ribeiro2018anchors}.
Post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure,
with the latter further classified into \emph{global} and \emph{local} methods
based on the locality in the input space~\cite{samek2021explaining}.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \vspace{-2.24cm}
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method)
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

@@

To address these limitations,
we propose R-LIME (Ruled LIME),
which provides both the contributions of each feature to the output
and the effective scope of the explanation.
R-LIME linearly approximates a complex decision boundary
in a rectangular region and maximizes the region
as long as the accuracy of the linear classifier exceeds a given threshold.
The region is interpretable for users because it is
expressed as a conjunction of feature predicates.
An example of the explanation by R-LIME for a sentiment prediction model
is shown in \cref{fig:example-rlime}.
It is clear that users can apply the insights derived from the explanation
only to the sentences containing the word ``not''.

@@

They are broadly divided into
\emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Model-dependent methods,
such as deep Taylor decomposition (DTD)~\cite{montavon2017explaining}
and layer-wise relevance propagation (LRP)~\cite{bach2015pixel},
primarily focus on neural networks and
explain the model's behavior using its parameters~\cite{samek2021explaining}.
While these methods provide detailed explanations
(e.g., layer-wise explanations for neural networks),
it is often challenging
to apply the same method to models with different structures.
In contrast,
model-agnostic methods use only the model's output.
Although they are applicable to any model,
they cannot explain the internal reasoning processes of the model.

Model-agnostic methods are further categorized into
\emph{global} and \emph{local} methods based on their locality in the input space.
Global methods,
such as partial dependence plots (PDP)~\cite{friedman2001greedy}
and accumulated local effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as \hl{individual conditional expectation (ICE)}~\cite{goldstein2015peeking},
local interpretable model-agnostic explanations
(LIME)~\cite{ribeiro2016why}, Anchor~\cite{ribeiro2018anchors} and
\hl{shapley additive explanations (SHAP)}~\cite{lundberg2017unified},
explain the model's behavior in the vicinity of a specific input.
While they offer explanations simpler and more accurate than global methods,
the scope of the explanation is limited locally.

@@

\subsection{Previous Work}
We specifically focus on \emph{local} and \emph{model-agnostic} methods.
This section briefly reviews existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
using a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation process involves the following steps:
\begin{enumerate}
  \item Generate a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and their pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) Convert $x$ into a binary vector $x'\in{\{0,1\}}^{m'}$,
        (ii) generate perturbed samples by randomly drawing non-zero elements
        from $x'$,
        and (iii) convert the perturbed samples back to the original space.
  \item Train a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function that assigns larger weights
        for samples closer to $x$, typically defined using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing each feature's contribution to the output $f(x)$.
However, it does not explicitly define the region for generating perturbed samples,
making it difficult for users to assess the effective scope of the explanation
~\cite{ribeiro2018anchors}.

@@

\subsection{Overview}
We propose R-LIME,
a novel method designed to address the limitations of LIME~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors}.
Similar to LIME,
it locally approximates the black-box classifier $f$
around the focal point $x$ using a linear classifier $g$,
and similar to Anchor,
it generates the perturbed samples for approximation
from a rectangular region (\cref{fig:rlime}).

@@

  \begin{algorithm}[p]
    \small
    \input{src/tex/alg4}
  \end{algorithm}
}
The R-LIME algorithm
is mainly based on the method used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main-problem},
greedy search are often used.
But greedy methods often converge to local optima,
so we use beam search, which selects multiple candidates at each iteration
to improve the search robustness.
The pseudo-code is shown in \cref{alg:greedy-search}.

@@

\subsubsection{Searching Rules with Highest Accuracy}
Given the set of candidate rules $\cands$,
the algorithm selects the $B$ candidate rules with the highest accuracy.
This problem can be formulated
as best arm identification in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered as an arm,
and reward of arm $a_i$ follows a Bernoulli distribution
with $P(X=1)=\Prec(A_i)$.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using $z$ and $f(z)$ after each trial.
To efficiently search the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm~\cite{kaufmann2013information}.
The pseudo-code is shown in \cref{alg:best-cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy the assumption.
This issue is discussed further in \cref{sec:reward}.

@@

\subsection{Computational Complexity}\label{subsec:complexity}
Post-hoc explanation methods including LIME, Anchor and R-LIME
need to sample a perturbation vector
and get the output of the black-box model multiple times,
which is computationally expensive.
The number of samples required for LIME is $|\mathcal{Z}_p|$,
which is the number of samples designated by the user.
On the other hand,
the expected number of samples required for Anchor and R-LIME is bounded by
$\mathcal{O}[m\cdot\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}]$,
where $\mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}$
is the expected number of samples
for best arm identification finding the best $B$ arms from $B\cdot m$ arms.
For the KL-LUCB algorithm~\cite{kaufmann2013information},
\begin{equation}
  \mathcal{O}_{\mathrm{MAB}[B\cdot m,B]}=
  \mathcal{O}\left[\frac{Bm}{\epsilon^2}\log\frac{Bm}{\epsilon^2\delta}\right].
\end{equation}
Then the total expected number of samples for Anchor and R-LIME is bounded by
\begin{equation}
  \mathcal{O}\left[\frac{Bm^2}{\epsilon^2}\log\frac{Bm}{\epsilon^2\delta}\right].
\end{equation}

For each iteration of the KL-LUCB algorithm, R-LIME needs to update
the linear classifier $g_i$, which is not required in Anchor.
If we use logistic regression as the linear classifier and update it
by stochastic gradient descent (SGD)~\cite{robbins1951stochastic},
the computational complexity of updating $g_i$ is $\mathcal{O}(m)$.
It is negligible compared to the complexity of generating a perturbed sample,
which is $\mathcal{O}(m^2)$ if we get a sample
from a multivariate normal distribution using Cholesky decomposition in advance.
Overall, the computational complexity of R-LIME is comparable to that of Anchor.

\section{Experiments}
To verify the effectiveness of our method,
we conducted qualitative and quantitative comparisons of R-LIME
with LIME and Anchor, using a real-world dataset.
\hl{Our code for R-LIME is available on GitHub}\footnote{\url{https://github.com/g-ohara/rlime}}.

@@

We used the recidivism dataset~\cite{schmidt1988predicting} for our experiments.
The dataset contains personal information on 9549 prisoners released from
North Carolina prisons between July 1, 1979 and June 30, 1980.
As shown in \cref{tab:rcdv}, the dataset includes 19 items such as
race (\emph{Race}), gender (\emph{Gender}),
presence of alcohol dependence (\emph{Alcohol}),
number of prior offenses (\emph{Priors}),
and presence of recidivism (\emph{Recidivism}).
For this experiment,
we treated the binary classification problem of predicting
the presence of recidivism (\emph{Recidivism}) as the target label.
We discretized continuous features and removed missing values,
resulting in 15 features.

We splitted the dataset into training data (7639 instances) and test data (955 instances),
and trained a random forest model with 50 trees as the black-box classifier
using the training data.
Then, we generate LIME, Anchor and R-LIME explanations
for two instances extracted from the test data (\cref{fig:instance}).
For R-LIME, we used logistic regression as the linear approximation model,
and a multivariate normal distribution estimated from the training data
as the distribution $\mathcal{D}$.
For both Anchor and R-LIME, the beam width was set to $B=10$,
the confidence coefficient to $1-\delta=0.95$,
and the tolerance of the KL-LUCB algorithm to $\epsilon=0.05$.
The accuracy threshold $\tau$ was set to $\tau=0.70,0.90$.

@@

\subsection{\hl{Quantitative Evaluation: Anchor vs. R-LIME}}\label{sec:exp-anchor}
\subsubsection{\hl{Experimental Setup}}
\hl{%
  To demonstrate that R-LIME explanations are more general than Anchor,
  we conducted a comparison of the coverage of explanations
  between Anchor and R-LIME\@.
  Under the same settings as in
}\cref{sec:exp-setting},
\hl{%
  we generated Anchor and R-LIME explanations
  for 704 instances from the test data of the recidivism dataset,
  under the values of $\tau=0.65,0.70,0.75,0.80,0.85,0.90$.
}

@@

\section{Conclusion}
Existing methods for
local model-agnostic explanations of black-box classifiers,
such as LIME and Anchor,
have limitations that they cannot achieve interpretability of
both the explanation and its application scope.
To address these challenges,
we proposed R-LIME,
a method that locally and linearly approximates the decision boundary
of a black-box classifier and provides a rectangular approximation region,
which is interpretable for users due to being expressed as a conjunction of feature predicates.
We proposed an algorithm to
maximize coverage of the approximation region
as long as the accuracy of the linear approximation model exceeds a given threshold.
Comparing R-LIME with existing methods on the real-world dataset,
we demonstrated that R-LIME achieves interpretability of both the explanation
and its application scope,
and provides explanations more accurate than LIME and more general than Anchor.
Finally,
we discussed the instability of behavior with imbalanced label distributions,
raised questions about the theoretical validity of using the KL-LUCB algorithm,
and hyperparameter tuning in practice.
